### 2024-05-02 thesis/chapters/03-design.tex
Complete design

The separability of fault groups into clusters is qualitatively analyzed by projecting normalized complete feature sets onto the plain utilizing two principal components. The transformed TD and FD sets are shown in Figure~\ref{fig:design:pca-complete-sets} with original labels and after low severities are assigned ``normal'' class. The decision boundaries that color the feature space around the data points are determined in 0.02x0.02 square areas by k-NN with five neighbours.

@@

However, the groups are not as clear-cut for frequency-domain features. Overall chaos in frequency spectra can be attributed to PC1: \emph{flux, entropy, negentropy, noisiness}, and the shape of frequency distribution to PC2: \emph{roll-on, roll-off}.

\section{Accelerometer data logger}
% Block diagram of device

% HW parameters and mention temporary device
\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                           & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                                  & STMicroelectronics \\ \hline
\textbf{Bus}                                     & SPI                \\ \hline
\textbf{Axis}                                    & 1 or 3             \\ \hline
\textbf{Range} (g)                               & $\pm$ 2 - 16      \\ \hline
\textbf{Bandwidth} (kHz)                          & 5 - 6.3            \\ \hline
\textbf{Sensitivity} (mg/LSB)                    & 0.06 - 0.49       \\ \hline
\textbf{Noise density} ($\mu g / \sqrt{\mathrm{Hz}}$ rms) & 75                 \\ \hline
\textbf{Microcontroller}                         & ESP32-PoE-ISO      \\ \hline
\textbf{CPU SoC}                                 & ESP32-WROOM-32     \\ \hline
\textbf{Output data rate} (kHz)                  & 26.7               \\ \hline
\textbf{ADC resolution} (bit)                    & 16                 \\ \hline
\textbf{FIFO}                                    & 3 kB (512 samples) \\ \hline
\end{tabular}
\caption{Hardware parameters of accelerometer data logger}
\label{tab:design:hw-sensors}
\end{table}

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{assets/design/hw-block-schematic.png}
	\caption{Sensor unit hardware block diagram}
	\label{fig:design:block-diagram-hw}
\end{figure}

% Task communication sequence diagram

% Firmware
% Timing is crutial 19ms.
Activity diagrams describe three basic actions of firmware for logging accelerometer samples:

\begin{itemize}
\itemsep0pt
\item begin new recording, 
\item log block of samples into CSV file,
\item finish recording.
\end{itemize}

\begin{figure}[h]
    \centering
	\includegraphics[width=\textwidth]{assets/design/firmware-design.png}
\end{figure}

@@

\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.17\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/machine-fan.jpg}
        \caption{\footnotesize Fan}
        \label{fig:machine:fan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.25\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/copeland-compressor.jpg}
        \caption{\footnotesize Compressor}
        \label{fig:machine:compressor}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/ksb-pump.jpg}
        \caption{\footnotesize Water pump KSB}
        \label{fig:machine:pump-ksb}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.31\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/sigma-pump.jpg}
        \caption{\footnotesize Water pump Sigma}
        \label{fig:machine:pump-sigma}
    \end{subfigure}
    \caption{Machines dedicated for vibration measurements}
\end{figure}

\textbf{Standing fan} is model \emph{Kalorik~TKG~VT1037} (Fig.~\ref{fig:machine:fan}) and one unit is available to us. It serves as a test bed during the sensor unit development. The sensor is placed on the plastic casing at the back of the drive motor. The fan has a 45~cm diameter with 3~propellers and a power of 45~W (class~\rom{1}). It has a switch for 3 rotational speeds, which are approximately 18.5~Hz (1100~rpm), 20.4~Hz (1200~rpm), and 22~Hz (1300~rpm).

\textbf{Scroll compressor} is model \emph{Copeland~ZR16} (Fig.~\ref{fig:machine:compressor}). Altogether, two units are available in two independent air conditioning units for the data center. The compressor has 9.7~kW of power (class~\rom{1}) and rotates at 2900 rpm (48.3 Hz). Two possible measurement locations are located on the sides on top of the bearings, just above the base and above the scroll.

\textbf{Water pumps} are available as 3 units in municipal drinking water pumping station. The apparatus consists of a single-stage axially split volute casing pump and an attached electric induction motor.

The newer primary pumps with bundled wireless cloud monitoring system are two units of \emph{KSB Omega~300-560} (Fig.~\ref{fig:machine:pump-ksb}). The pumps were commissioned in 2018, and they rotate at 1493 rpm (24.9 Hz). The electric motor provides 400~kW of power (class~\rom{3}).

The secondary pump is one unit named \emph{Sigma~300-OVD-600} (Fig.~\ref{fig:machine:pump-sigma}) installed in 1986. It rotates at 1485 rpm (24.75 Hz), and its electric motor has power of 450~kW (class~\rom{3}). Each pump and motor have 2 bearings. Therefore, there are 4 measurement positions in total.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-rms.png}
        \caption{Vibration rms velocity dashboard}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-spectrum.png}
        \caption{Vibration spectral analysis dashboard}
    \end{subfigure}
     \caption{KSB Guard cloud monitoring for pumps}
     \label{fig:design:ksb-guard}
\end{figure}

\textbf{Complete schedule} for planned vibration data gathering on the designated machines and measurement procedure is described in Appendix~\ref{appendix:technical-docs}. In addition, we are able to export the entire historical record at hourly intervals. Logs from \textbf{KSB Guard} cloud monitoring tool contain vibration rms velocities and frequency spectra for two monitored KSB water pumps (Figure~\ref{fig:design:ksb-guard}).

\section{Data volume savings}
The apparent advantage of feature discovery is reducing the amount of data downstream. Data compression must occur on edge devices to enable the utilization of wireless low-power wide area networks (LPWAN). The protocol stack may differ, so goodput is compared without node configuration metadata and keepalive messages.

The machinery monitoring system relies on determining several parameters:
\begin{itemize}
\itemsep0pt
\item \textbf{Number of source channels} ($S$): comprises the number of monitored machines, measurement locations for sensors, and active sensor axes.
\item \textbf{Sampling frequency} ($f_s$): is set based on the linear response of the accelerometer, the types of faults intended for detection, and how soon they should be noticed after they arise. The higher required sensitivity means a higher sampling rate derived according to the Nyquist theorem. At a minimum, it should be 15 kHz to 20 kHz.
\item \textbf{Interval between successive measurements} ($T$): specifies the minimal response time to sudden failure. The more critical the machine is, the interval should be shorter. The bigger the machine parts, the slower the defect evolves.
\item \textbf{Duration of valid recording} ($D$): is the captured snapshot of machine unaltered behavior associated with a timestamp. Duration should cover at least 3 windows for spectral estimation. The spectral resolution of 1 Hz amounts to 3 seconds of signal under such assumptions.
\item \textbf{Number of extracted features} ($F$): are ideally key trend indicators pointing to symptoms of common malfunctions. We aim for a total of 6 features.
\end{itemize}

@@

As an example to approximate required network goodput and storage in practice, we consider continuous vibration \textbf{monitoring for municipal water pumping station}. The station has 3 pumps and 3 electric motors.

A pump and motor pair have 4 bearings together for drive end and non-drive end positions. Each position has a sensor mounted in 3 directions that makes a total of \emph{36 source channels}. The sampling frequency at each position is set to \emph{20 kHz}. The recordings have \emph{duration of 5 seconds} and are triggered regularly every 1 hour (\emph{8760 times per year}).

In a year, the system gathers 31.54 Gs (gigasamples) which is 58.74 GiB with a 16-bit ADC resolution. Reasonably precise spectral estimation with 10 thousand bins needs 3.15 Gs per year. On the other hand, 6 features out of each channel keep only 1.89 Ms per year for a lossy compression ratio of 16667:1. Low data volumes potentially enable feature selection and models to be offloaded directly to edge devices. The entire machine's history can be kept in a small flash memory module.

\section{Data collection methodology}
% UML of experiment

% Measurement plan (calendar and folders) - dataset schema

>>>>

The separability of fault groups into clusters is qualitatively analyzed by projecting normalized complete feature sets onto the plain utilizing two principal components. The transformed TD and FD sets are shown in Figure~\ref{fig:design:pca-complete-sets} with original labels and after low severities are assigned ``normal'' class. The decision boundaries that color the feature space around the data points are determined in 0.02x0.02 square areas by k-NN with five neighbours. We observe the inability to separate faults by linear boundaries in given feature spaces and the noncompactness of clusters.

@@

However, the groups are not as clear-cut for frequency-domain features. Overall chaos in frequency spectra can be attributed to PC1: \emph{flux, entropy, negentropy, noisiness}, and the shape of frequency distribution to PC2: \emph{roll-on, roll-off}. PCA efficiently expresses attributes in less dimensional space, but the resulting linear combination is hard to comprehend for explaining decisions.

\section{Accelerometer data logger}
Data acquisition of vibrations from industrial machinery necessitates the assembly of the standalone data logger. The required hardware modules of the device include a sensor as a triaxial MEMS accelerometer, a fast persistent memory unit, and a microcontroller to control communication with both peripherals. The recording starts with a button press, and the indicator LED provides feedback about the operation in progress.

More detailed specification according to standards creates demands especially on the accelerometer, in terms of bandwidth for linear response to at least 1 kHz, ideally more. The sampling rate should be greater than 2 kHz for each of the three axes (6 kSps). Out of the options available on the market, we opted for the accelerometers with digital communication bus with more than 12-bit sample resolution, high sensitivity, and low noise density.

\begin{table}[h]
\renewcommand{\arraystretch}{1.2}
\centering
\begin{tabular}{|l|l|l|}
\hline
\textbf{Accelerometer}                   & \textbf{IIS3DWB}   \\ \hline
\textbf{Vendor}                          & STMicroelectronics \\ \hline
\textbf{Bus}                             & SPI  (to 10 MHz)               \\ \hline
\textbf{Axis}                            & 1 or 3             \\ \hline
\textbf{Range}                           & $\pm$ 2 - 16 g (19 - 157 $\mathrm{m/s}^2$)   \\ \hline
\textbf{Bandwidth}                       & 5 - 6.3 kHz           \\ \hline
\textbf{Sensitivity}                     & 0.06 - 0.49 mg/LSB      \\ \hline
\textbf{Noise density} & 75 $\mu \mathrm{g} / \sqrt{\mathrm{Hz}}$ rms                \\ \hline
\textbf{Output data rate}               & 26.8 kHz             \\ \hline
\textbf{Sample resolution}              & 16 bit                 \\ \hline
\textbf{FIFO}                           & 3 kB (512 samples) \\ \hline
\textbf{Microcontroller}                & \textbf{ESP32-PoE-ISO}      \\ \hline
\textbf{CPU SoC}                                 & ESP32-WROOM-32     \\ \hline
\textbf{CPU clock rate}                          & 80 MHz     \\ \hline
\end{tabular}
\caption{Hardware parameters of accelerometer data logger}
\label{tab:design:hw-sensors}
\end{table}

The hardware configuration we arrived at for the data logger powered with a power bank via USB connector is in Table~\ref{tab:design:hw-sensors}. The IIS3DWB accelerometer is a cheap enough MEMS accelerometer that approaches industrial standards for vibration monitoring. The evaluation board for the accelerometer STEVAL-MKI208V1K is connected via a ribbon cable that enables proper attachment to designated places.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{assets/design/hw-block-schematic.png}
	\caption{Data logger hardware block diagram}
	\label{fig:design:block-diagram-hw}
\end{figure}

ESP32-PoE-ISO is chosen as a microcontroller development kit because it has an SD card slot connected via an SD/MMC bus. It enables us to store all samples from FIFO. The accelerometer uses an SPI bus with a maximum speed of 10 MHz that can be connected to any physical GPIO pin. The block diagram of the designed hardware device is in Figure~\ref{fig:design:block-diagram-hw}.

\begin{figure}[h]
    \centering
	\includegraphics[width=\textwidth]{assets/design/firmware-design.png}
	\caption{Activity diagram of data logger firmware functions}
	\label{fig:design:firmware-activity}
\end{figure}

The firmware's role lies in coordinating hardware adapters and straightforward vibration acquisition for later analysis. The provided functionality is hence minimalistic. Two user-driven actions, one periodic event, and one timer-generated event satisfy these basic requirements. Activity diagrams outline the steps of the firmware actions~(Fig.~\ref{fig:design:firmware-activity}).

@@

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-1.png}
        \caption{Fan speed 1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-2.png}
        \caption{Fan speed 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-3.png}
        \caption{Fan speed 3}
    \end{subfigure}
    \caption{Frequency spectrum of audio recorded in close proximity to the side of the standing fan}
    \label{fig:design:fan-speed}
\end{figure}

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/standing-fan.jpg}
        \caption{\footnotesize Fan}
        \label{fig:machine:fan}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/compressor.jpg}
        \caption{\footnotesize Compressor}
        \label{fig:machine:compressor}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/ksb-pump.jpg}
        \caption{\footnotesize Water pump KSB}
        \label{fig:machine:pump-ksb}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/sensor/sigma-pump.jpg}
        \caption{\footnotesize Water pump Sigma}
        \label{fig:machine:pump-sigma}
    \end{subfigure}
    \caption{Machines dedicated for vibration measurements}
    \label{fig:design:machines}
\end{figure}

\textbf{Standing fan} is model \emph{Kalorik~TKG~VT1037} (Fig.~\ref{fig:machine:fan}) and one unit is available to us. It serves as a test bed during the data logger development. The sensor is placed on the plastic casing at the back of the drive motor. The fan has a 45~cm diameter with 3~propellers and a power of 45~W (class~\rom{1}). It has a switch for 3 rotational speeds, which are approximately 19~Hz, 21~Hz, and 23~Hz ($\pm$ 0.18 Hz). Speed estimation was done by spectral analysis of audio (Fig.~\ref{fig:design:fan-speed}) in \emph{Spectroid} app and confirmed by 240 fps high-speed camera.

\textbf{Scroll compressor} is model \emph{Copeland~ZR16} (Fig.~\ref{fig:machine:compressor}). Altogether, two units are available in two independent air conditioning units for the data center. The compressor has 9.7~kW of power (class~\rom{1}) and rotates at 2900 rpm (48.3 Hz). Possible measurement locations are located on the sides of the compressor atop the bearings, just above the base and below the scroll. Steel casing is not in direct contact with the bearing which causes significant alteration to the signal.

\textbf{Water pumps} are available as three units in municipal drinking water pumping station. The apparatus consists of a single-stage axially split volute casing pump and an attached electric induction motor. The newer primary pumps with bundled wireless cloud monitoring are two units of \emph{KSB Omega~300-560} (Fig.~\ref{fig:machine:pump-ksb}). The pumps were commissioned in 2018, and they rotate at 1493 rpm (24.9 Hz). The motor \emph{WEG W50} provides 400~kW of power (class~\rom{3}). The bearing designations at numbered places for defect frequency  calculations are 6319-C3 (1), 6324-C3 (2), and 6317-2Z (3 \& 4)

The secondary pump is one unit named \emph{Sigma~300-OVD-600} (Fig.~\ref{fig:machine:pump-sigma}) installed in 1986. It rotates at 1485 rpm (24.75 Hz), and its electric motor has a power of 450~kW (class~\rom{3}). Each pump and motor have two bearings, hence there are four measurement positions.

\section{Data collection methodology}
Vibration measurement for one placement on the machine involves three trials (two for Sigma pump). After each trial, the sensor is detached and attached again to introduce ``noise''. The sensor is mounted to the machine on a flat surface with adhesive of thin double-sided carpet tape.

Placements are numbered for simplicity (Fig.~\ref{fig:design:machines}) but the precise description follows the notation of MIMOSA convention from ISO~13373-1. The sensor is mounted on compressor's placem: \emph{SFTA001AT000TN} (1), \emph{SFTA002AT000TN} (2), on WEG motor's positions: \emph{MTRA001AT000TN} (1), \emph{MTRA002AT045TN} (2), and on KSB pump positions: \emph{PMPA003AT000TN} (3), \emph{PMPA004AT000TN} (4).

\begin{table}[]
\begin{tabular}{|l|c|c|ccccc|}
\hline
\multirow{2}{*}{\textbf{Machine}}           & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Facility code}}} & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Label}}} & \multicolumn{5}{l|}{\textbf{Date}}                                                                                                                                                                                   \\ \cline{4-8} 
                                            & \multicolumn{1}{l|}{}                                           & \multicolumn{1}{l|}{}                                & \multicolumn{1}{l|}{\textbf{20/02/2024}} & \multicolumn{1}{l|}{\textbf{27/02/2024}} & \multicolumn{1}{l|}{\textbf{05/03/2024}} & \multicolumn{1}{l|}{\textbf{19/03/2024}} & \multicolumn{1}{l|}{\textbf{26/03/2024}} \\ \hline
\multirow{2}{*}{\textbf{Scroll compressor}} & K3                                                              & C1                                                   & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{$\bigtimes$}         &                                          \\ \cline{2-8} 
                                            & K5                                                              & C2                                                   & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{$\bigtimes$}         &                                          \\ \hline
\multirow{2}{*}{\textbf{Water pump (KSB)}}  & no. 1                                                           & P1                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \cline{2-8} 
                                            & no. 7                                                           & P2                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \hline
\multirow{2}{*}{\textbf{Pump's motor}}      & no. 1                                                           & M1                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \cline{2-8} 
                                            & no. 7                                                           & M2                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{$\bigtimes$}         & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \hline
\textbf{Pump (Sigma)}                       & no. 5                                                           & P3                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \hline
\textbf{Motor (Sigma)}                      & no. 5                                                           & M3                                                   & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & \multicolumn{1}{c|}{}                    & $\bigtimes$                              \\ \hline
\end{tabular}
\caption{Data collection schedule}
\label{fig:design:data-schedule}
\end{table}

The triaxial recording has a duration 60~s (at $f_s$ = 26.8~kHz) and a 16-bit resolution. One file has 24.6~MiB as a binary stream and 48~MiB in TSV format. 
Measurements were executed and organized into folders according to schedule in Table~\ref{fig:design:data-schedule}. It should be noted that because of operational regulations, the second KSB pump was measured the day after the first. The estimated total space requirements for all 92 uncompressed raw recordings is 2.21~GiB and approximately 4.31~GiB as TSV files.

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-rms.png}
        \caption{Vibration rms velocity dashboard}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
    		\centering
        \includegraphics[width=\textwidth]{assets/design/ksb-guard-spectrum.png}
        \caption{Vibration frequency analysis dashboard}
    \end{subfigure}
     \caption{KSB Guard cloud monitoring for pumps}
     \label{fig:design:ksb-guard}
\end{figure}

Logs from \textbf{KSB Guard} cloud monitoring tool contain vibration rms velocities and frequency spectra for two monitored KSB water pumps (Figure~\ref{fig:design:ksb-guard}). We were able to export the historical log of vibration velocity for last year at hourly intervals and sample frequency spectra.

\section{Data volume savings}
The apparent advantage of feature discovery is reducing the amount of data downstream. Data compression must occur on edge devices to enable the utilization of wireless low-power wide area networks (LPWAN). The protocol stack may differ, so goodput in this section is compared without node configuration metadata and keepalive messages.

The machinery monitoring system relies on determining a few parameters:
\begin{itemize}
\itemsep0pt
\item \textbf{Number of source channels} ($S$) - comprises the number of monitored machines, measurement locations for sensors, and active sensor axes.
\item \textbf{Sampling frequency} ($f_s$) - is set based on the linear response of the accelerometer, the types of faults intended for detection, and how soon they should be noticed after they arise. The higher required sensitivity means a higher sampling rate derived according to the Nyquist theorem. At a minimum, it should be 15 kHz to 20 kHz.
\item \textbf{Interval between successive measurements} ($T$) - specifies the minimal response time to sudden failure. The more critical the machine is, the shorter the interval should be. The bigger the machine parts, the slower the defect evolves.
\item \textbf{Duration of valid recording} ($D$) - is the captured snapshot of machine unaltered behavior associated with a timestamp. Duration should cover at least twelve windows for spectral estimation.
\item \textbf{Number of extracted features} ($F$) - are ideally key trend indicators pointing to symptoms of common malfunctions. We aim for a total of six features.
\end{itemize}

@@

As an example to approximate required network goodput and storage in practice, we consider continuous vibration \textbf{monitoring for municipal water pumping station}. The station has three pumps and three electric motors.

A pump and motor pair have four bearings together for drive end and non-drive end positions. Each position has a sensor mounted in three directions that makes a total of \emph{36 source channels}. The sampling frequency at each position is set to \emph{20 kHz}. The recordings have \emph{duration of five seconds} and are triggered regularly every 1 hour (\emph{8760 times per year}).

In a year, the system gathers 31.54 Gs (gigasamples) which is 58.74 GiB with a 16-bit ADC resolution. Reasonably precise spectral estimation with 10 thousand bins needs 3.15 Gs per year. On the other hand, 6 features out of each channel keep only 1.89 Ms per year for a lossy compression ratio of 16667:1. Low data volumes potentially enable feature selection and models to be offloaded directly to edge devices. The entire history of the machine's health could be stored in a small flash memory module.

### 2024-05-02 thesis/chapters/05-evaluation.tex
Analysis finished - page 37
Complete design

\chapter{Evaluation} \label{section:evaluation}

@@

\section{Validation of data logger}
% Validation of firmware on standing fan - speeds
\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-1.png}
        \caption{Speed 1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-2.png}
        \caption{Speed 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-3.png}
        \caption{Speed 3}
    \end{subfigure}
    \caption{Frequency spectrum from sound of standing fan}
\end{figure}

>>>>

\chapter{Evaluation} \label{chapter:evaluation}

@@

\subsection{Data logger validation}

### 2024-05-02 thesis/chapters/TEXT.tex
Complete design

\subsection{Statistical tests}
The statistical tests and visual checks are conducted to assess \textbf{normality and stationary} of time series. Half a second of amplitude samples are used from every sensor channel. These 25 thousand observations are downsampled tenfold to $2500$.

\textbf{Shapiro-Wilk's test} rejects the null hypothesis (${p < 0.05}$) that data is drawn from a normal distribution under most circumstances. The signal has normal distribution when it resembles pink noise lacking a regular pattern or weak exhibition of fault symptoms. \textbf{Quantile–quantile plots} confirm non-normal distribution because of the striking samples tilt to the diagonal line.

\textbf{Augmented Dickey-Fuller test} rejects the null hypothesis of unit root (${p < 0.001}$). The same is confirmed with the \textbf{autocorrelation} function shape. It denotes that the stochastic process is stationary as the oscillation is bounded.

@@

The Euclidean norm of feature in the triaxial vector eliminates reliance on the direction of measurement. Value ranges are depicted in Figure~\ref{fig:design:feature-range}.

%\begin{figure}[ht]
%    \centering
%    \begin{subfigure}[b]{\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/feature-range-temporal.png}
%        \caption{Time-domain features}
%    \end{subfigure}
%    \hfill
%    \begin{subfigure}[b]{\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/feature-range-spectral.png}
%        \caption{Frequency-domain features}
%    \end{subfigure}
%    \caption{Feature value ranges in inner bearing position (A)}
%    \label{fig:design:feature-range}
%\end{figure}

Fault labels and their severities come from the directory structure within the dataset. The binary target variable indicating whether to initiate a warning is called an anomaly. Anomalies are labeled according to relative fault severity level. We decided to investigate two fault severities having levels above 0.6 and 0.9. The quantity of observation by fault and anomaly severity is shown in Table \ref{tab:observation-counts}. The dataset is substantially unbalanced.

Pearson's correlation of features to rpm is very low in the whole dataset. In the temporal domain, the correlation coefficient is within an interval of -0.08 to 0.26. In the spectral domain, the correlation to rpm for all FFT window sizes from $2^8$ up to $2^{14}$ is mostly very low from -0.14 up to 0.26, except for the centroid being around 0.35.

 temporal domain that are present in these pairs (ordered from the most correlated): \{impulse, clearance\}, \{crest, impulse\}, \{crest, clearance\}, \{pp, rms\}. In the spectral domain set \{skewness, kurtosis\} has a strong correlation.

%\begin{figure}[h!]
%    \centering
%    \includegraphics[width=0.8\textwidth]{assets/design/pca-explained-variance.png}
%    \caption{Number of principal components to cumulative explained variance percentage in the inner bearing position}
%    \label{fig:design:pca-explained-variance} 
%\end{figure}

The variables are also more inter-correlated in the temporal domain shown by principal component analysis. For 95\% of the explained variance of PCA 3 components (98.69\%) are needed in the temporal domain whereas 4 in the spectral domain (95.26\%). Figure \ref{fig:design:pca-explained-variance} visualizes cumulative explained variance with an increasing number of principal components.

%\begin{figure}[h!]
%    \centering
%    \begin{subfigure}[b]{0.49\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/pca-loading-plot-temporal.png}
%        \caption{Temporal features}
%    \end{subfigure}
%    \hfill
%    \begin{subfigure}[b]{0.49\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/pca-loading-plot-spectral.png}
%        \caption{Spectral features}
%    \end{subfigure}
%    \caption{PCA loading plots for min-max scaled features from the inner bearing position}
%    \label{fig:design:pca-loading-plot} 
%\end{figure}

PCA efficiently expresses attributes in less dimensional space, but the resulting linear combination is hard to comprehend for explaining decisions.

@@

Lowering the number of attributes allows visualization of labels in planar scatter plots. The data points in 3-dimensional cross-sections are colored by true fault classes of all severities in Figure \ref{fig:design:feature-space-scatter}. We observe the inability to separate faults by linear boundaries in given feature spaces and the noncompactness of clusters.

Cluster overlaps of various defect types are also confirmed by silhouette scores around zero. Therefore, the k hyperparameter cannot be increased much, as it would cause the merge of mutually confused categories. Low-severity faults should be relabeled as a class without fault to watch out just for alert conditions.

%\begin{figure}[ht]
%    \centering
%    \begin{subfigure}[b]{\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/best-features-scatter-fault-A-temporal.png}
%        \caption{Temporal domain features}
%    \end{subfigure}
%    \hfill
%    \begin{subfigure}[b]{\textwidth}
%        \includegraphics[width=\textwidth]{assets/design/best-features-scatter-fault-A-spectral.png}
%        \caption{Spectral domain features}
%    \end{subfigure} 
%    \caption{Cross sections of 3-dimensional feature spaces with ground truth labels for best attributes. Features are chosen by rank product method.}
%    \label{fig:design:feature-space-scatter}
%\end{figure}

@@

We suspected that the PCA property of maximizing variances in each of the few dimensions would support more separation of clustered groups and would be beneficial in classification. The bar chart supports this assumption. PCA always produces greater model accuracy than any other method of feature selection. The rank product of the three scoring techniques is not better every time, but it balances out aggregated methods to achieve more stable results across testing sets.

In every instance, a subset of spectral features is responsible for better accuracy than a subset of temporal features. Presumably, it is so because of many correlated pairs of attributes in the temporal domain. Further tests should be made to determine optimal feature subsets and the impact of altering the feature election process.

@@

\section{Sensor hardware and drivers}
In preliminary exploratory vibration measurements of machines, the ADXL335 accelerometer is connected to BeagleBone Black microcontroller. The bandwidth, sensitivity, and sampling frequency of the makeshift solution do not appear to be adequate (Table~\ref{tab:design:hw-sensors}).

The IIS3DWB accelerometer is a cheap enough MEMS accelerometer that approaches industrial standards for vibration monitoring. ESP32-PoE-ISO is chosen as a microcontroller development kit because it has an SD card slot connected via SD/MMC bus. It enables us to store all samples from FIFO. The accelerometer uses an SPI bus with a maximum speed of 10 MHz that can be connected to any physical GPIO pin. The block diagram of the designed hardware device is in Figure~\ref{fig:design:block-diagram-hw}.

>>>>

@@

@@

@@

@@

\section{Sensor hardware and drivers}
